---
title: "Untitled"
author: "Astrid Langsrud, Håkon Gryvill"
date: "25 4 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Keywords

Issues to be addressed:

# Introduction
Describe the problem you want to study.
Why is this interesting?
What prior knowledge do you have?
What do you want to achieve?

In our experiment we want to study how different factors affect our responsiveness. 
The experiment was performed in the following way: 
A ruler is dropped at a random point in time. The test person's responsiveness is determined by how far down the ruler is caught.
This experiment is of interest because it gives us an idea of how mental and physical distraction affects our responsiveness. (Skriv mer svada)



# Selection of factors and levels:
Which factors do you think are relevant to the problem described above?
Do you expect an interaction between some of the factors?
Which levels should be used, and why do you think these are reasonable?
How can you control that the factors really are at the desired level?

We are looking at the following factors in our experiment: the gender of the test person, physical distraction (i. e., whether the person has been spinning around or not) and mental distraction (the person has to talk about a given subject during the experiment).
In advance we expected that at least physical distraction would decrease the responsiveness. 
However, we did not expect that any of the factors would have a major effect on each other. For instance, there is no reason for us to think that spinning around affects women more than men. 

# Selection of response variable:
Which response variable will provide information about the problem described above?
Are there several response variables of interest?
How should the response be measured?
What can you say about the accuracy of these measurements?

(We have chosen how far down the ruler is caught as response variable, because this gives an idea of the response time.)
We found it difficult to make accurate measurements of the response variable. The response variable was measured by taking the average of the numbers covered by the fingers (dårlig formulering?).  
# Choice of design:
2 k factorial.
2 k–p fractional factorial or other design?
Desired resolution of the design?
Is it necessary or desirable to use a blocked design?
Is it necessary or desirable with replicates?

In order to see significant effects, two replicates of the experiment was performed. In both replicates, all experiments were performed in random order.
# Implementation of the experiment:
Randomization.
Describe any problems with the implementation (maybe the randomization was not followed?).
Is each experiments a genuine run replicate, that is reflects the total variability of the experiment? (Each trial should be a performed independently and constitute a full trial.)
As we see it, there are two main problems with the performance of the experiment. Firstly, it is almost impossible to make two genuine run replicates. Our responsiveness improves as we keep doing the experiments. Secondly, the experiments are not completely independent. Spinning around might affect the responsiveness longer than expected. 
# Analysis of data

Calculation of effects and assessment of statistical significance.
Check the assumptions. Important: residual plots!
```{r, echo=TRUE, eval = TRUE}

y1 = c(16, 8, 15, 14, 15, 20, 11, 0)
y2 = c(13, 11, 14, 12, 20, 23, 14, 21)
y = c(y1, y2)
library(FrF2)
plan <- FrF2(nruns=8, nfactors=3, replications = 2, randomize=FALSE)
plan <- add.response(plan,y)
plan
lm3 <- lm(y~.^3,data=plan)
cubePlot(lm3,"A","B","C")
MEPlot(lm3) # main effects plot
IAPlot(lm3) # interaction effect plots
effects <- 2*lm3$coeff
effects
```

# Conclusion and recommendations:
Which conclusions can you draw from the experiment?
Interpretation of significant effects, main and interaction plots.
Remember that plots are illustrative and very useful for demonstrations.

